What We Eval in the Shadows

### 2021-04-17 paper/main.tex
sec 1 to 5

\mypara{Kaggle.} The Kaggle website ({\small \url{kaggle.com}}) is an online
platform for data science. It allows users to submit problems and compete to
solve them. Solutions, called \emph{kernels}, are uploaded as plain scripts or
notebooks. The quality of code on Kaggle is not uniform. Each kernel is
runnable, and thus there is one script per kernel. Input data is provided. We
obtained \KaggleKernels kernels. Since Kaggle is not curated, we used SHA-1 hashes
to identify and remove \KaggleDuplicates duplicate entries. The remaining
\KaggleUnique kernels are unique solutions to \KaggleCompetitions competitions.
They contain \KaggleCode lines of R code. Only \KaggleWithEvals kernels call
\eval.

@@

\medskip
\begin{compactenum}
\item \emph{Download.} Packages are downloaded from CRAN. For scripts, a web
  crawler retrieves code, and the Kaggle command-line tool gets data.
  Installation is complicated by native dependencies, which are not adequately
  documented and thus hard to resolve automatically.
\item \emph{Extract.} Given installed packages and scripts, the next step is to
  create runnable programs. The \genthat tool helpfully extracts all runnable
  code snippets from a package and turns each of these into a self-standing
  program~\cite{issta18}. Some Kaggle kernels are already scripts. For those,
  nothing more needs to be done. Others kernels are packages as notebooks, either   as Rmarkdown or Jupyter. For those, we use \c{knitr} to extract
   runnable code.
  The body of each extracted program is instrumented with calls to our dynamic
  analyzer to ensure that we only record calls to \eval from the code of
  interest and not from bootstrapping or execution harness operations.
\item \emph{Trace.} Each program executes using our dynamic analysis tool, a heavily instrumented interpreter that captures calls to \eval and
  many other fine-grained runtime events. Packages are run twice, once to
  capture \eval calls originating from package code and a second time to
  capture calls coming from the base libraries. To avoid any interference, each
  program is run in its own process.
  As the GNU R bytecode compiler is written in R, we turn it off
  to avoid recording \eval in its code, and fallback on the interpreter.
\item \emph{Analyze.} Finally, the analysis output is merged, cleaned, and summarized
  in a post-processing phase driven by series of R scripts. The summarized data
  is then analyzed in RMarkdown notebooks to gather insights. All figures and
  numbers appearing in the paper are generated automatically. Figures are
  produced in PDF by \c{ggplot2}; numbers are exported as \LaTeX macros.
\end{compactenum}

\medskip\noindent The code extraction and tracing steps of the pipeline are run
in parallel~\cite{GNUparallel} orchestrated by a Makefile. Servers have
identical environments thanks to docker images with all dependencies installed.

@@

The dynamic analysis is performed by \rdyntrace, a modified R virtual machine
based on GNU R 4.0.2 that exposes low-level callbacks for a variety of runtime
events~\cite{oopsla19b}.

The tracer registers callbacks to all \eval functions as well as a few other
ones that allow us to identify better where the expressions come from. For
example, we taint the results of calls to \c{parse} (turns strings into
expressions) and \c{match.call} (reflects on current call). We also capture calls
to R API for dynamic code loading (\eg\xspace \c{library}, \c{require}, and
\c{source}). Next, we subscribe to the events related to a variable definition
and assignment, allowing us to record side effects that happen in environments
while evaluating code in \eval. One challenge we encountered during the analysis
was that essentially R only provides source references for top-level calls in a
function body or in the block surrounded by braces. That is to say, a method
whose body would \c{eval(x)} would not have debug information that we can use to
identify that particular \eval. The same expression surrounded by braces would
have a line number and source reference. This is unfortunately not easily fixed
in the R implementation. Instead, we extended the dynamic analysis tool to
attach synthetic source code references to all \eval call sites by traversing
ASTs. The tracer is implemented as an R package in 3.2K C++ and 1.3K of R code.
For performance reasons, most of the tracing is done in C++. While in theory, the
implementation is relatively straightforward, it is not so in practice:
%
\begin{compactitem}[---]

@@

\item While the R interpreter is implemented in C, a lot of the core
  functionality is done in R. For example, package loading is implemented in R
  using \eval or the S4 and R6 object systems. This makes it hard in the tracer
  to separate the \eval that are essential to user programs from the
  accidental ones that are products of the way R implements its basic
  operations. This is even more so for the side-effects analysis.

@@

\mypara{Limitations.} The analysis has some limitations. Even with the extension
described above, there are still \PkgUndefinedRnd \eval calls without source
references. This occurs when a reference to the \eval function is passed as an 
argument to a higher-order function or when the \eval call originates from
native code. However, these missing source references account for a meager
\PkgUndefinedRatio of all calls; they are unlikely to affect our results. Other
limitations are that we ignore calls to the native \eval function and to the
alternate \c{rlang::tidy\_eval} function, which uses native \eval internally.
The \c{rlang} package introduces a new kind of promise called \c{quosure} which
is evaluated by \c{tidy\_eval}. None of these limitations should invalidate our
conclusions.

@@

\begin{wrapfigure}{r}{5cm} \hspace*{-12mm}
  \centering
  \includegraphics[width=72mm]{pkgs-eval-callsites-hist.pdf} \caption{CRAN
  \eval call sites}%
  \label{fig:pkgs-eval-callsites-hist}
\end{wrapfigure}
%
This section focuses on CRAN packages and reports statistics about the usage of
\eval. We use the word \emph{site} to an occurrence of a call site to the \eval
function in the source code and \emph{call} to denote an observed invocation of
the \eval function.

\subsection{CRAN}

There are \PkgEvalCallSites \eval sites in \PkgPackages packages. The proportion
of packages calling \eval is \PkgPackagesRatio. Over half of these packages
have fewer than 3 sites, and with the exception \MaxEvalCallSitesPackage, which
has \MaxEvalCallSitesCount sites, all packages contain fewer than
\MaxEvalCallSitesRest sites (\cf Fig~\ref{fig:pkgs-eval-callsites-hist}, which
shows a histogram of sites per package). These sites appear in \PkgFunsWithEval
functions (\CranFunsWithEvalRatio of all functions in CRAN).

\begin{figure}[!b]
\small
\begin{tabular}{@{}l@{\hspace{1.5cm}}l@{}}
\begin{minipage} {5cm}
  \begin{tabular}{|r@{\,}r@{\,}l@{}r|r@{\,}r@{\,}l@{}r|} \hline
    \multicolumn{3}{|c}{\small\#calls} &\small \#pck
&     \multicolumn{3}{c}{\small\#calls} &\small\#pck \\\hline
\tt 1 &--& \tt 10      & \packageBina  & \tt 10K &--&\tt 100K  & \packageBine\\
\tt 11 &--& \tt 100    & \packageBinb  & \tt 100K &--&\tt 1M  & \packageBinf\\
\tt 101 &--& \tt 1K    & \packageBinc  & \tt 1M &--&\tt 10M   & \packageBing\\
\tt 1K &--& \tt 10K    & \packageBind  & \tt 10M &--& \tt 100M & \packageBinh\\\hline
\end{tabular}
\caption{Call frequency}\label{freq}
\end{minipage}
&
\begin{minipage}{7cm}
\begin{tabular}{|@{\,}r|rrrr|}\hline
  &\eval & \c{evalq} & \c{eval} & \c{local}\\[-1.5mm]
           & & & \c{.parent} &\\\hline
\small Static sites &\packageStaticeval&\packageStaticevalq&\packageStaticevalparent&\packageStaticlocal \\
\small Exercised sites&\packageTriggeredeval&\packageTriggeredevalq&\packageTriggeredevalparent&\packageTriggeredlocal\\
\small Invocations&\packageEvalsRnd&\packageEvalqsRnd&\packageEparentsRnd&\packageLocalsRnd\\\hline
\end{tabular}~\\[2mm]\caption{Variants}\label{tab:variantseval}
\end{minipage}\end{tabular}

%\end{figure}

%\begin{figure}[b] \centering
  \includegraphics[width=.78\textwidth]{traced-eval-callsites.pdf} \centering
  \caption{\eval call sites coverage of the \PkgPackages packages.}%
  \label{fig:traced-eval-callsites}
\end{figure}

For dynamic analysis, we run \CranRunnableScripts programs extracted from
\CranPackages packages. Any run that does not exercise \eval is discarded. This
left \packageNbruns runs from \packageCorpus packages. There were
\packageAllcalls calls in \packageTriggeredpkgs packages originating from
\PkgHitEvalCallSites unique sites. In terms of coverage, the data exercised
\PkgHitEvalCallSitesAvgRatio of sites, a coverage similar to the code coverage
metric for the packages, which is \PkgCodeCoverage. The fact that not all sites
are exercised can be chalked down to incomplete tests and occasional analysis
failures (\PkgFailedProgramsRatio of programs crashed or timed out).
Fig~\ref{fig:traced-eval-callsites} shows the number of sites that were
exercised; coverage is unequal. Figure~\ref{freq} summarizes the frequency of
dynamic calls to \eval, with the left column being the number of calls and the
right, the number of packages in that range. There are \packageFewcalls packages
with low \eval frequency, fewer than 100 calls, and a smaller, but still
significant, number of packages, \packageManycalls to be precise, that use \eval
more than 1,000 times. Package \packageMaxcallspack makes \packageMaxcalls calls
and thus accounts for over half of the observations.
Figure~\ref{tab:variantseval} summarizes the use of variants of \eval. For each
of the four variants, the first row shows the number of sites in the corpus
(\emph{static}), the second row is the number of sites that were encountered
during analysis (\emph{exercised}), the last row is the number of calls
(\emph{invocations}). The overwhelming majority of sites and calls are to \eval
itself, \c{eval.parent} is rare, and both \c{evalq} and \c{local} are barely
used at all. The difference between sites and exercised sites underscores the
limitations due to code coverage.

@@

\begin{tabular}{c}
  \vspace*{-1mm}
  {\hspace{-25mm}\includegraphics[width=0.7\textwidth]{package_size_loaded_distribution}}
\end{tabular}
\caption{Loaded code} \label{fig:sizedistribution}
\end{wrapfigure}

@@

\Eval takes any value, but if its argument is not an expression, \eval returns
it unchanged. Expressions account for \packageCodepercent of arguments in our
corpus. More specifically, \packageSymbolpercent are \texttt{symbol}s (single
variables such as \c{x}), \packageLanguagepercent are \c{language} objects
(function calls such as \c{f(x)} or \c{x+1}), and \packageExpressionpercent are
\c{expression} objects (lists of expressions such as function calls or
symbols). Further inspection reveals that most symbols,
\packageGgplotsymbolpercent to be exact, come from a single site in the
\c{ggplot2} package and have the value \c{\_inherit}.\footnote{This models
inheritance in \c{ggproto}, one of the many object-oriented systems in
R, used exclusively in the \c{ggplot2} graphics library.}

@@

\begin{figure}[tb!]
\begin{tabular}{@{}c@{}c@{}}
\begin{minipage}{7.5cm}
 \includegraphics[width=\textwidth]{package_events_per_pack_small}
\end{minipage}&\begin{minipage}{7.5cm}
  \includegraphics[width=\textwidth]{package_events_per_pack_large}
\end{minipage}\\[-3mm]
\small (a) Small & \small (b) Large
\end{tabular}
 \caption{Instructions per call} \label{ev}
\end{figure}

\subsection{Base}\label{sec:usage-base}

We recorded \baseAllcalls \eval calls in the \BasePackages base libraries. This
comes from running 10\% randomly selected programs from the
\CranRunnableScriptsRnd extracted programs from CRAN packages, covering
\baseTriggeredevalpct of the \BaseEvalCallSites sites. Most of the calls
(\baseEvalsratio) are to the \c{eval} function, and most arguments
(\baseCodepercent) are expressions. However, unlike CRAN, the majority
(\baseLanguagepercent) of the arguments are language objects. The median
argument size is \baseMedianszeval, which is more than for CRAN and the maximum size is \baseMaxszeval, less than for CRAN. Small instruction counts ($<50$) amount to \baseSmalleventspct of calls. A single site in the \c{match.arg}
function is responsible for \baseTopFuncPercent of all recorded calls. This
function provides a convenient way for argument verification using partial
matching. It is heavily relied on in R functions that use string arguments to
parameterize function behavior, a popular API mechanism in R. For example, in a
body of \c{center <- function(x, type = c("mean", "median", "trimmed"))} one
can get the value of \c{type} parameter using \c{type <- match.arg(type)}. A
call \c{center(x, "trim")} will match \c{type} to \c{"trimmed"}.

The challenge dealing with Base is that every program uses it. It can thus
generate extreme amounts of data, but the data is quite predictable. For the
rest of the paper, we only mention Base when there are surprising
observations.

\medskip

\subsection{Kaggle}

In total, \kaggleAllcalls \eval were recorded, all to \eval function. Out of
\kaggleStaticeval sites from \KaggleWithEvals scripts, only \kaggleTriggeredeval sites
in \kaggleNbruns scripts were hit; \KaggleFailedScripts scripts failed to run,
and the others did not exercise the \eval sites. This is partially expected
as Kaggle code does not need to abide by any checks. Upon a manual inspection,
we observed that indeed, the failing scripts were of poor quality, often not
finished, using misspelled package names, hard-coded file paths, or accessing
missing files.

Most call sites are invoked only once. Only one site is called \kaggleMaxcalls times.
Expressions amount for \kaggleCodepercent of arguments. Unlike the CRAN corpus,
there are very few symbols (\kaggleSymbolpercent). Most expressions result from
calls to \c{parse}; thus, most \evals start with strings. The median argument
size is \kaggleMedianszeval, which makes sense, as few arguments are only
symbols. The largest argument is \kaggleMaxszeval. The distribution of
instructions per eval is similar to CRAN; \kaggleSmalleventspct of \evals
execute fewer than 50 instructions. Manual inspection of \eval usage in Kaggle
suggests that it is consistent with the data obtained for CRAN. We do not
discuss it further.

\medskip

\subsection{Discussion}
These results show that \eval is central to the language implementation as it is
omnipresent in the relatively small Base library. The Base library is, in turn,
used by every single package. To avoid having Base dominate the result, we have
excluded calls occurring there from the data we report when analyzing packages.
CRAN Packages, which are typically developed by experienced programmers, make
regular and varied use of \eval. They represent our most interesting data set as
these packages result from over 20 years of contributions by thousands of
authors. The code in packages is well maintained and relatively well tested.
Finally, Kaggle scripts are often written by less sophisticated users and
likely perform more straightforward tasks, thus having a lesser need for \eval, and its uses originate from strings as these are easier to manipulate for end-users. Due to the relative paucity of data in Kaggle, we do not pursue the data set further.

@@

\medskip\noindent\EE{$min(e)=\c{V}$} Expressions that represent values are
frequently passed to \eval -- they occur in 17\% of sites. The majority of
those, \packageValOneNodePercent, are inline constants (integer or double
vectors). The rest trivially evaluate to a value, \eg~\c{1+1}.\footnote{True as
long as base functions such as \c{+} are not redefined. They typically are not,
but this is a limitation of this categorization.} In our corpus,
\packageNbCallSitesUniqueActualValue call sites only ever see a simple value.
Manual inspection reveals cases such as the following
\begin{lstlisting}
 f1 <- eval(paste("A~",paste(paste(names(X[,-1])),collapse="+")))
\end{lstlisting}
where the argument to \eval is a string which \eval simply returns. Of twenty
randomly selected \evals, \packageUsefulValueEvalPercent of them  need % There are a majority which needs eval
\eval for some inputs. Either we are dealing with a value that needs to be
constructed dynamically, or the value is a default case that sometimes is
replaced by a more interesting expression. This form is usually evaluated in few
interpreter steps; in fact, the median is only
\packageMinimizedmedianoperationscRnd. The environment in which they evaluate is
mostly irrelevant (unless a built-in operator is redefined).

@@

\ \\[5mm] % To push the box on the next page
\medskip\noindent\EE{$min(e)=\c{F()}$} This form captures simple function calls,
\eg~\c{f(2)}, with neither variables nor assignments. More specifically, it
allows for variables in the function position but not in arguments. Usually,
looking up function names does not trigger computation, but that is not a given
if the function is returned by a promise or if the function name is shadowed by
a promise containing a value, then the computation will occur. This form occurs in
6\% of sites and typically does not perform much work in the interpreter.

>>>>

\mypara{Kaggle.} The Kaggle website ({\small \url{kaggle.com}}) is an online
platform for data science. It allows users to submit and compete to solve
problems. Solutions, called \emph{kernels}, are uploaded as scripts or
notebooks. While the quality of code is not uniform, each kernel is runnable,
and maps to ones script. Input data is provided. We obtained \KaggleKernels
kernels. Since Kaggle is not curated, we used SHA-1 hashes to identify and
remove \KaggleDuplicates duplicate entries. The remaining \KaggleUnique kernels
are unique solutions to \KaggleCompetitions competitions. They contain
665K lines of R code. Only \KaggleWithEvals kernels call \eval.

@@

\medskip
\begin{compactenum}
\item \emph{Download.} Packages are downloaded from CRAN. For scripts, a web
  crawler retrieves code, and the Kaggle command-line tool gets data.
  Installation is complicated by native dependencies, which are not adequately
  documented and thus hard to resolve automatically.
\item \emph{Extract.} Given installed packages, the \genthat tool extracts all
  runnable code snippets and turns each of these into a self-standing
  program~\cite{issta18}. Some Kaggle kernels are already scripts and nothing
  more needs to be done. Kernels released as notebooks are processed by
  \c{knitr} which extracts runnable code. The body of each program is
  instrumented with calls to our dynamic analyzer to ensure that we only record
  calls to \eval from the code of interest and not from bootstrapping or
  execution harness operations.
\item \emph{Trace.} Each program executes using our dynamic analysis tool, a
  heavily instrumented interpreter that captures calls to \eval and many other
  fine-grained runtime events. Packages are run twice, once to capture \eval
  calls originating from package code and a second time to capture calls coming
  from the base libraries. To avoid any interference, each program is run in its
  own process with the GNU R compiler turned off to avoid recording its
  execution.
\item \emph{Analyze.} Finally, the analysis output is merged, cleaned, and summarized
  in a post-processing phase driven by series of R scripts. The summarized data
  is then analyzed in RMarkdown notebooks to gather insights. All figures and
  numbers appearing in the paper are generated automatically. Figures are
  produced in PDF by \c{ggplot2}; numbers are exported as \LaTeX macros.
\end{compactenum}

\noindent The pipeline runs in parallel~\cite{GNUparallel} orchestrated
by a Makefile. Servers have identical environments thanks to docker images with
all dependencies installed.

@@

The dynamic analysis is performed by \rdyntrace, a modified R virtual machine
based on GNU R 4.0.2 that exposes low-level callbacks for a variety of runtime
events~\cite{oopsla19b}. The tracer registers callbacks to all \eval variants
and a few additional functions to assist in locating the origin of \eval
arguments. For example, we taint the results of calls to \c{parse} and
\c{match.call} (reflects on current call). We also capture calls to R API for
dynamic code loading. Next, we subscribe to the events related to a variable
definition and assignment, allowing us to record side effects that happen in
environments while evaluating code in \eval. One challenge was that R only
provides source references for block surrounded by braces. Thus, a function
whose body is \c{eval(x)} would no usable debug information, while the same
expression surrounded by braces would be fine. This is unfortunately not easily
fixed. We extended the dynamic analysis tool to attach synthetic source code
references to all \eval call sites by traversing ASTs. The tracer is implemented
as an R package in 3.2K C++ and 1.3K of R code. For performance reasons, most of
the tracing is done in C++. While in theory, the implementation is relatively
straightforward, not so in practice:
%
\begin{compactitem}[---]

@@

\item While the R interpreter is implemented in C, most core functionality is in
  R. For example, package loading is implemented in R using \eval. This makes it
  hard in the tracer to separate the \eval that are essential to user programs
  from the accidental ones that are products of the way R implements its basic
  operations. This is even more so for the side-effects analysis.

@@

\mypara{Limitations.} Even with the extension described above, there are
\PkgUndefinedRnd \eval calls without source references. This occurs when \eval
is passed as an argument to a higher-order function or when the \eval call
originates from native code. However, these missing source references account
for a meager \PkgUndefinedRatio of all calls; they are unlikely to affect our
results. Other limitations are that we ignore calls to the native \eval and to
the alternate \c{rlang::tidy\_eval} uses native \eval internally. None of these
limitations should invalidate our conclusions.

@@

\begin{wrapfigure}{r}{5cm} \vspace{-1cm} \hspace*{-12mm}
  \centering
  \includegraphics[width=72mm]{pkgs-eval-callsites-hist.pdf} \caption{CRAN
  \eval call sites}%
  \label{fig:pkgs-eval-callsites-hist}\vspace{-3mm}
\end{wrapfigure}
%
This section focuses on CRAN packages and reports statistics about the usage of
\eval. We use the word \emph{site} to an occurrence of a call site to the \eval
function in the source code and \emph{call} to denote an observed invocation of
the \eval function.

\begin{figure}[!b]
\small
\begin{tabular}{@{}l@{\hspace{1.5cm}}l@{}}
\begin{minipage} {5cm}\small
  \begin{tabular}{|r@{\,}r@{\,}l@{}r|r@{\,}r@{\,}l@{}r|} \hline
    \multicolumn{3}{|c}{\small\#calls} &\small \#pck
&     \multicolumn{3}{c}{\small\#calls} &\small\#pck \\\hline
\tt 1 &--& \tt 10      & \packageBina  & \tt 10K &--&\tt 100K  & \packageBine\\
\tt 11 &--& \tt 100    & \packageBinb  & \tt 100K &--&\tt 1M  & \packageBinf\\
\tt 101 &--& \tt 1K    & \packageBinc  & \tt 1M &--&\tt 10M   & \packageBing\\
\tt 1K &--& \tt 10K    & \packageBind  & \tt 10M &--& \tt 100M & \packageBinh\\\hline
\end{tabular}
\caption{Call frequency}\label{freq}
\end{minipage}
&
\begin{minipage}{7cm}\small
\begin{tabular}{|@{\,}r|rrrr|}\hline
  &\eval & \c{evalq} & \c{eval} & \c{local}\\[-1.5mm]
           & & & \c{.parent} &\\\hline
\small Static sites &\packageStaticeval&\packageStaticevalq&\packageStaticevalparent&\packageStaticlocal \\
\small Exercised sites&\packageTriggeredeval&\packageTriggeredevalq&\packageTriggeredevalparent&\packageTriggeredlocal\\
\small Invocations&\packageEvalsRnd&\packageEvalqsRnd&\packageEparentsRnd&\packageLocalsRnd\\\hline
\end{tabular}~\\[2mm]\caption{Variants}\label{tab:variantseval}
\end{minipage}\end{tabular}

\end{figure}

\begin{figure}[b] \centering
  \includegraphics[width=.7\textwidth]{traced-eval-callsites.pdf} \centering
  \caption{\eval call sites coverage of the \PkgPackages packages.}%
  \label{fig:traced-eval-callsites}
\end{figure}

\paragraph{CRAN}

@@

\begin{tabular}{c}
  \vspace*{-1mm}
  {\hspace{-25mm}\includegraphics[width=0.7\textwidth]{package_size_loaded_distribution}}
\end{tabular}
\caption{Loaded code} \label{fig:sizedistribution}
\end{wrapfigure}
There are \PkgEvalCallSites \eval sites in \PkgPackages packages. The proportion
of packages calling \eval is \PkgPackagesRatio. Over half of these packages
have fewer than 3 sites, and with the exception \MaxEvalCallSitesPackage, which
has \MaxEvalCallSitesCount sites, all packages contain fewer than
\MaxEvalCallSitesRest sites (\cf Fig~\ref{fig:pkgs-eval-callsites-hist}, which
shows a histogram of sites per package). These sites appear in \PkgFunsWithEval
functions (\CranFunsWithEvalRatio of all functions in CRAN).
For dynamic analysis, we run \CranRunnableScripts programs extracted from
\CranPackages packages. Any run that does not exercise \eval is discarded. This
left \packageNbruns runs from \packageCorpus packages. There were
\packageAllcalls calls in \packageTriggeredpkgs packages originating from
\PkgHitEvalCallSites unique sites. In terms of coverage, the data exercised
\PkgHitEvalCallSitesAvgRatio of sites, a coverage similar to the code coverage
metric for the packages, which is \PkgCodeCoverage. The fact that not all sites
are exercised can be chalked down to incomplete tests and occasional analysis
failures (\PkgFailedProgramsRatio of programs crashed or timed out).
Fig~\ref{fig:traced-eval-callsites} shows the number of sites that were
exercised; coverage is unequal. Figure~\ref{freq} summarizes the frequency of
dynamic calls to \eval, with the left
column being the number of calls and the right, the number of packages in that
range. There are \packageFewcalls packages with low \eval frequency, fewer than
100 calls, and a smaller, but still significant, number of packages,
\packageManycalls to be precise, that use \eval more than 1,000 times. Package
\packageMaxcallspack makes \packageMaxcalls calls and thus accounts for over
half of the observations. Figure~\ref{tab:variantseval} summarizes the use of
variants of \eval. For each of the four variants, the first row shows the number
of sites in the corpus (\emph{static}), the second row is the number of sites
that were encountered during analysis (\emph{exercised}), the last row is the
number of calls (\emph{invocations}). The overwhelming majority of sites and
calls are to \eval itself, \c{eval.parent} is rare, and both \c{evalq} and
\c{local} are barely used at all. The difference between sites and exercised
sites underscores the limits of code coverage.

@@

\Eval takes any value, but if its argument is not an expression, it is returned
unchanged. Expressions account for \packageCodepercent of arguments. More
specifically, \packageSymbolpercent are \texttt{symbol}s (single variables such
as \c{x}), \packageLanguagepercent are \c{language} objects (function calls),
  and \packageExpressionpercent are \c{expression} objects (lists of
  expressions). Further inspection reveals that most symbols,
  \packageGgplotsymbolpercent to be exact, come from a single site in the
  \c{ggplot2} package and have the value \c{\_inherit}.\footnote{This models
  inheritance in \c{ggproto}, one of the many object-oriented systems in R, used
  in the \c{ggplot2} graphics library.}

@@

\begin{figure}[h!]
\begin{tabular}{@{}c@{}c@{}}
\begin{minipage}{7.5cm}
 \includegraphics[width=\textwidth]{package_events_per_pack_small}
\end{minipage}&\begin{minipage}{7.5cm}
  \includegraphics[width=\textwidth]{package_events_per_pack_large}
\end{minipage}\\[-3mm]
\small (a) Small & \small (b) Large
\end{tabular}
 \caption{Instructions per call} \label{ev}
\end{figure}

\paragraph{Base}\label{sec:usage-base}

We recorded \baseAllcalls \eval calls in the \BasePackages base libraries. This
comes from running 10\% randomly selected programs from the
\CranRunnableScriptsRnd extracted programs from CRAN packages, covering
\baseTriggeredevalpct of the \BaseEvalCallSites sites. Most of the calls
(\baseEvalsratio) are to the \c{eval} function, and most arguments
(\baseCodepercent) are expressions. However, unlike CRAN, the majority
(\baseLanguagepercent) of the arguments are language objects. The median
argument size is \baseMedianszeval, which is more than for CRAN and the maximum
size is \baseMaxszeval, less than for CRAN. Small instruction counts ($<50$)
amount to \baseSmalleventspct of calls. A single site in the \c{match.arg}
function is responsible for \baseTopFuncPercent of all recorded calls. This
function provides a convenient way for argument verification using partial
matching. It is heavily relied on in R functions that use string arguments to
parameterize function behavior, common practice in R. For example, in a body of
\c{center<-function(x,type=c("mean","median","trimmed"))} one can get the
value of \c{type} parameter using \c{type<-match.arg(type)}. A call
\c{center(x,"trim")} will match \c{type} to \c{"trimmed"}. The challenge
dealing with Base is that every program uses it. It can thus generate extreme
amounts of data, but the data is quite predictable. From now on, we only mention
Base when it surprises.

\paragraph{Kaggle}

In total, \kaggleAllcalls \eval were recorded, all to \eval function. Out of
\kaggleStaticeval sites from \KaggleWithEvals scripts, only \kaggleTriggeredeval
sites in \kaggleNbruns scripts were hit; \KaggleFailedScripts scripts failed to
run, and the others did not exercise the \eval sites. This is partially expected
as Kaggle code does not need to abide by any checks. Upon a manual inspection,
we observed that indeed, the failing scripts were of poor quality, often not
finished, using misspelled package names, hard-coded file paths, or accessing
missing files. Most call sites are invoked once. Only one site is called
\kaggleMaxcalls times. Expressions amount for \kaggleCodepercent of arguments,
there are very few symbols (\kaggleSymbolpercent). Most expressions result from
calls to \c{parse}; thus, most \evals start with strings. The median argument
size is \kaggleMedianszeval, which makes sense, as few arguments are only
symbols. The largest argument is \kaggleMaxszeval. The distribution of
instructions per eval is similar to CRAN; \kaggleSmalleventspct of \evals
execute fewer than 50 instructions. Manual inspection of \eval usage in Kaggle
suggests that it is consistent with the data obtained for CRAN. We do not
discuss it further.

\paragraph{Discussion}
\Eval is central to R's implementation as it is omnipresent in the Base library.
That library is, in turn, used by every single package. To avoid Base dominating
the result, we excluded calls to it from the results. CRAN Packages, which are
typically developed by experienced programmers, make regular and varied use of
\eval. They represent our most interesting data set as these packages result
from over 20 years of contributions by thousands of authors. The code in
packages is well maintained and relatively well tested. Finally, Kaggle scripts
are often written by less sophisticated users and likely perform more
straightforward tasks, thus having a lesser need for \eval, and its uses
originate from strings as these are easier to manipulate for end-users.

@@

\medskip\noindent\EE{$min(e)=\c{V}$} Expressions that represent values are
frequently passed to \eval -- they occur in 17\% of sites. The majority of
those, \packageValOneNodePercent, are inline constants (integer or double
vectors). The rest trivially evaluate to a value, \eg~\c{1+1}.\footnote{True as
long as base functions such as \c{+} are not redefined. They typically are not,
but this is a limitation of this categorization.} In our corpus,
\packageNbCallSitesUniqueActualValue call sites only ever see a simple value,
such as:
\begin{lstlisting}
 f1 <- eval(paste("A~",paste(paste(names(X[,-1])),collapse="+")))
\end{lstlisting}
where the argument to \eval is a string which \eval simply returns. Of twenty
randomly selected \evals, \packageUsefulValueEvalPercent of them need \eval for
some inputs. Either we are dealing with a value that needs to be constructed
dynamically, or the value is a default case that sometimes is replaced by a more
interesting expression. This form is usually evaluated in few interpreter steps;
in fact, the median is only \packageMinimizedmedianoperationscRnd. The
environment in which they evaluate is mostly irrelevant.

@@

\medskip\noindent\EE{$min(e)=\c{F()}$} This form captures simple function calls,
\eg~\c{f(2)}, with neither variables nor assignments. More specifically, it
allows for variables in the function position but not in arguments. Usually,
looking up function names does not trigger computation, but that is not a given
if the function is returned by a promise or if the function name is shadowed by
a promise containing a value, then the computation will occur. This form occurs in
6\% of sites and typically does not perform much work in the interpreter.
